Online Social Cohesion Reflects Real-World Group Action in Syria during the Arab Spring

### 2020-06-01 manuscript.Rmd
Editing intro

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and flyers generated on mimeograph machines,
through in-person meetings for social organizations and in churches, and through
word of mouth. They gathered together and hand-made protest signs that reflected
their agendas. As McAdam found, people were more likely to join movements when
they had a personal connection [@McAdam_1986; @McAdam_1986]. Strong ties were
better predictors of involvement in high-risk protest activity. However, the era
of the Internet and online social mobilization through decentralized networks
has upended theories about recruitment to, and participation in, social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@Threlkeld_2017]. The research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to big data analysis of language on social media. The
cohesiveness and strength of social movements in the pre-Internet era could be
operationalized by the number of people turning out to protest, and through the
political events that precipitated and resulted from social gatherings en masse.

In the offline era of social mobilization, the content of placards and protest
signs provided the protest narrative for onlookers near and far. For cohesive
social movements, the content of placards and posters itself was cohesive. For
example, the “I am a man” posters held by the men involved in the 1968
sanitation workers’ strike reflected their internal cohesion and unified,
goal-oriented movement [@Windsor_2020]. Social media replicates and amplifies
this phenomenon during moments of social unrest. Amid the cacophony of messages
during political upheaval, such as the Arab Spring, consistent themes emerge
through hashtags, retweets, and the repeated mentions of concepts such as
regime, security, demonstration, and human rights. The cohesiveness of these
themes, i.e., whether movement participants are “singing in one voice,” reflects
the real-world events happening contemporaneously with collective action in
social movements or civil conflicts.

In the present research, we investigate the relationship between linguistic
cohesion and social cohesion in times of social conflict and unrest. We tackle a
question related to the one posed by Zeitzoff (2017): how does social media
reflect on-the-ground realities during times of social conflict
[@Zeitzoff_2017]? We examine patterns of linguistic cohesion on Twitter during
the Arab Spring uprising in Syria between March and June 2012. We contextualize
the linguistic cohesion with social mobilization and response to the relative
intensity of the conflict. Cohesion is a multi-layered concept referring to the
connectedness of words, thoughts, speakers, and actions. Cohesion reflects the
continuity of language and expression, and it shows the connectivity of
participants in a social movement. To address the relationship between language
and discourse and social movements, we adopt an interdisciplinary approach that
incorporates linguistics, psychology, and political science using nonlinear
dynamical systems methodology to model the relationship between language and
action. We find that in the early months of 2012, conflict-related events were
more strongly associated with linguistic cohesion than were more neutral or
peaceful events.

Disentangling the complicated relationship between mass online mobilization, and
social mobilization in person, continues to be a relevant question. The
political processes that transpired during and after the Arab Spring are still
in the process of being understood. As Zeitzoff (2017) recognizes, social media
is changing conflict, but the mechanisms and relationships are still being
understood [@Zeitzoff_2017].

>>>>

Prior to the advent of social media, people coordinated their protest behavior
using phone tree calling sheets and mimeographed flyers,
through in-person meetings of social organizations and churches, and through
word of mouth. They gathered together and hand-made protest signs that reflected
their agendas. As McAdam found, people were more likely to join movements when
they had a personal connection [@mcadam1986recruitment; @mcadam1993specifying]: 
Historically, strong ties have been
strong predictors of involvement in high-risk protest activity. 
However, the rise of the Internet and its decentralized online networks
has upended theories about recruitment to---and participation in---social
movements. Strong interpersonal ties have yielded to weak impersonal ties that
strongly predict involvement in social movements [@steinert2017spontaneous]. Research
methodologies for measuring social movements have similarly evolved from
ethnographic interviews to ``big data'' analyses of language on social media. 
*(Probably need to wrap up with a sentence here.)*

The cohesiveness and strength of social movements in the pre-Internet era could be
operationalized by the number of people turning out to protest and through the
political events that precipitated and resulted from social gatherings *en masse*.
The content of placards and protest
signs provided the protest narrative in their own words. Cohesive
social movements were reflected in their cohesive placards and posters. For
example, the ``I am a man'' posters held by the men involved in the 1968
sanitation workers' strike reflected their internal cohesion and unified,
goal-oriented movement [@Windsor2020].

In the Internet age, social media replicates and amplifies
this phenomenon during moments of social unrest. Amid the cacophony of messages
during political upheaval---such as the Arab Spring---consistent themes emerge
through hashtags, retweets, and the repeated mentions of concepts such as
regime, security, demonstration, and human rights. The cohesiveness of these
themes (i.e., whether movement participants are metaphorically singing in one voice) 
reflects the real-world events happening contemporaneously with collective action in
social movements or civil conflicts.

In the present research, we investigate the relationship between linguistic
cohesion and social cohesion in times of social conflict and unrest. We tackle a
question related to the one posed by Zeitzoff [-@zeitzoff2017social]: How does 
social media reflect on-the-ground realities during times of social conflict?
To do so, we examine patterns of linguistic cohesion on Twitter during
the Arab Spring uprising in Syria between March and June 2012.

We contextualize linguistic cohesion with social mobilization and response to the relative
intensity of the conflict. Cohesion is a multi-layered concept referring to the
connectedness of words, thoughts, speakers, and actions. Cohesion reflects the
continuity of language and expression, and it shows the connectivity of
participants in a social movement. To address the relationship between language
and discourse and social movements, we adopt an interdisciplinary approach that
incorporates linguistics, psychology, and political science using nonlinear
dynamical systems methodology to model the relationship between language and
action.

Disentangling the complicated relationship between mass online mobilization---and
social mobilization in person---continues to be relevant, even as our modes
of communication change. The
political processes that transpired during and after the Arab Spring are still
in the process of being understood. As Zeitzoff [-@zeitzoff2017social] recognizes, 
social media is changing conflict, but the mechanisms and relationships by
which it does so are still being understood.
